Linear vertical shear instability in astrophysical disks: a semi-numerical study

### 2014-12-19 linear_adia.tex
added beta cooling requirements

\subsubsection{Qualitative expectations in the presence of thermal
  relaxation}

>>>>

\subsubsection{Qualitative expectations in the presence of thermal
  relaxation}
We have shown that when $\gamma>1$ so that there is a positive
vertical entropy gradient and perturbations evolve
adiabatically ($t_c\to\infty$) that the VSI is strongly
surpressed. This is the limit where the term proportional to
$t_c^{-1}$ in the linearized energy equation (Eq. \ref{lin_energy})
becomes negligible compared to the term proportional to
$\gamma/\Gamma-1$. The latter term is associated with the vertical
entropy gradient, or buoyancy frequency $N_z$.

Reversing the argument, we expect to recover the VSI when the bouyancy
frequency becomes negligible compared to the cooling rate, or 
\begin{align}
  t_c \ll \frac{1}{N_z}. 
\end{align} 
Parametrizing $t_c = \beta\Omega_k^{-1}$ and inserting the expression
for $N_z$ appropriate for vertically isothermal disks, we expect
\begin{align}
  \beta \ll
  \frac{1}{H_\mathrm{iso}\sqrt{\frac{\gamma-1}{\gamma}}\left|\frac{\p\ln\rho}{\p
          z}\right|} 
\end{align}
is required to enable the VSI. This inequality is satisfied at the
midplane and as $|z|\to\infty$. It can be satisifed at all heights if
it is satisfied at the height of $\mathrm{max}|\p\ln\rho/\p z|$. 
Using the expression for the density field for vertically isothermal
disks (Eq. \ref{vertiso_eqm}), we expect 
\begin{align}
  \beta \ll \epsilon\frac{3^{3/2}}{2}\sqrt{\frac{\gamma}{\gamma-1}},
\end{align}
i.e. $\beta\ll O(\epsilon)$, is a sufficient condition for
bouyancy to be negligible.
